### 2018-04-09 response_letter.tex
Add response letter

\documentclass[11pt,a4]{article}

%%%%%%%%%%%%
% Language %
%%%%%%%%%%%%
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage[english]{babel}

%% For graphics
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{epstopdf} % Must be loaded right after graphicx

\usepackage[normalem]{ulem} % to use the \sout command

\usepackage[natbib=true,backend=biber,style=alphabetic,
firstinits=true,doi=false,eprint=false,isbn=false,url=false,texencoding=utf8,bibencoding=utf8]{biblatex}
%\AtBeginBibliography{\footnotesize} % To make bibliography smaller

  %Must be used with biber as backend
  %Execute with "biber filename"
  \addbibresource[location=remote]{https://www.dropbox.com/s/r00o9lw76cksxvj/Library.bib?dl=1}

% Fonts and symbols
\usepackage{amsfonts,amssymb,amsmath} % AMS fonts and symbols package

\begin{document}

\title{Response to the Associate Editor and Reviewers' comments on the 
paper ''Distributed Nonlinear Control Design using Separable Control 
Contraction Metrics''}

\author{Humberto Stein Shiromoto, and Ian R. Manchester}

\date{\today}

\maketitle

\printbibliography

\end{document}