### 2018-04-09 response_letter.tex
Start response_letter

\documentclass[11pt,a4]{article}

%%%%%%%%%%%%
% Language %
%%%%%%%%%%%%
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage[english]{babel}

%% For graphics
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{epstopdf} % Must be loaded right after graphicx

\usepackage[normalem]{ulem} % to use the \sout command

\usepackage[natbib=true,backend=biber,style=alphabetic,
firstinits=true,doi=false,eprint=false,isbn=false,url=false,texencoding=utf8,bibencoding=utf8]{biblatex}
%\AtBeginBibliography{\footnotesize} % To make bibliography smaller

  %Must be used with biber as backend
  %Execute with "biber filename"
  \addbibresource[location=remote]{https://www.dropbox.com/s/r00o9lw76cksxvj/Library.bib?dl=1}

% Fonts and symbols
\usepackage{amsfonts,amssymb,amsmath} % AMS fonts and symbols package

\begin{document}

\title{Response to the Associate Editor and Reviewers' comments on the 
paper ''Distributed Nonlinear Control Design using Separable Control 
Contraction Metrics''}

\author{Humberto Stein Shiromoto, and Ian R. Manchester}

\date{\today}

\maketitle

\section*{Associate Editor} % (fold)
\label{sec:ae}
All the reviewers found the paper clear and well written. However, the
main concerns that have been raised involve:

\begin{enumerate}
	\item novelty nature of the results - the authors should better 
	clarify the novelties of their work and better relate their results 
	with the literature;

	\item Effectiveness of the results - the authors
should provide some sufficient conditions for the existence of the
sum-separable metric. How this relates to the structure of the graph?
It seems that the matrix inequalities should be satisfied along any
possible $(x,u)$. If this is the case, then this seems to make the
resolution of the inequalities extremely challenging. The authors
should discuss this point;

	\item The results of this paper seem to provide a natural extension to
previous works by the authors. What are the benefits of this extension?
Are there cases that cannot be studied with the previous work by the
authors but can be studied with the approach of this paper?

\item Related to the items above, the authors should also include in the
paper a revised application section for e.g. a large scale network
system arising in engineering applications. Platooning of autonomous
vehicles would be, for example, an interesting application. This would
allow the authors to comment on the items raised above.
\end{enumerate}
% section  (end)

\section{Reviewer\# 6}\label{r:6}

\section{Reviewer\# 9}\label{r:9}

\section{Reviewer\# 11}\label{r:11}

\section{Reviewer\# 13}\label{r:13}

\printbibliography

\end{document}